### 2021-02-16 GALI_2020.Rmd
mixed model analysis

---
title: "Data Responsive-report"
author: "Ouafaa Hmaddi"
date: "January 18, 2021"
output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#load libraries
#if (!requireNamespace("BiocManager", quietly = TRUE))
#install.packages("caret")

#BiocManager::install("preprocessCore")

library(rio)
library(tidyverse)
library(lubridate)
library(janitor)
library(ggplot2)
library(readxl)
library(reshape2)
library(readr)
library(psych)
library(foreign)
library(plm)
library(lme4)
library(lattice)
library(lmerTest)
library(EFAutilities)
library(haven)
library(sjPlot)
library(XLConnect)
library(preprocessCore)
library(knitr)
library(caret)

```

```{r Countries to analyse in main sample}
gali <- import("Data/GALI_External_DataRelease_2020.xlsx")

count(gali,info_venture_country_hq)

#Removing countries with less than 6 ventures.

countries_to_keep <- gali %>% 
  group_by(gali$info_venture_country_hq) %>%
  summarize(Count=n()) %>%
  mutate(Percent = round((Count/sum(Count)*100))) %>%
  arrange(desc(Count))

#X countries have fewer than 70 firms in the sample. Because so few observations may not offer good representation, I excluded firms headquartered in these countries from the analysis, keeping X firms in Y countries and Z industries.

countries_to_keep <- filter(countries_to_keep, `gali$info_venture_country_hq` != 'NA')

countries_to_keep <- filter(countries_to_keep, Count >6)
countries_to_keep$`gali$info_venture_country_hq`

gali <- gali %>%
  filter(info_venture_country_hq  %in% countries_to_keep$`gali$info_venture_country_hq`)

#Number of ventures in main sample 
#kable(countries_to_keep)

```

```{r}
gali <- filter(gali, found_name1_age > 14 | is.na(found_name1_age))
gali <- filter(gali, found_name2_age > 14 | is.na(found_name2_age))
gali <- filter(gali, found_name3_age > 14 | is.na(found_name3_age))
gali <- filter(gali, found_name1_age < 70 | is.na(found_name1_age))
gali <- filter(gali, found_name2_age < 70 | is.na(found_name2_age))
gali <- filter(gali, found_name3_age < 70 | is.na(found_name3_age))

#Keep only ventures with 3 years of post acceleration information 
gali <- filter(gali, fu1report_followup_yes == 1 & fu2report_followup_yes == 1 & fu3report_followup_yes == 1)

#firms that are 5 years or younger when they entered the acceleration program (when data was collected)

if (gali$application_year == 2013) { gali <- filter(gali, info_founding_year > 2007)
}  else if (gali$application_year == 2014) {gali <- filter(gali, info_founding_year > 2008)
} else if (gali$application_year == 2015) {gali <- filter(gali, info_founding_year > 2009)
}  else if (gali$application_year == 2016) {gali <- filter(gali, info_founding_year > 2010)
} else if (gali$application_year == 2017) {gali <- filter(gali, info_founding_year > 2011)
}  else gali <- filter(gali, info_founding_year > 2012)
```

```{r Countries to keep in sample with valid data}

gali <- gali %>%
  rename(country = info_venture_country_hq) %>%
  rename(year = program_year)

gali$country[gali$country == "United States of America"] <-"USA"
gali$country[gali$country == "Tanzania, United Republic of"] <-"Tanzania"

countries_to_keep <- gali %>% 
  group_by(gali$country) %>%
  summarize(Count=n()) %>%
  mutate(Percent = round((Count/sum(Count)*100))) %>%
  arrange(desc(Count))

countries_to_keep <- filter(countries_to_keep, `gali$country` != 'NA')

countries_to_keep <- filter(countries_to_keep, Count >6)
countries_to_keep$`gali$country`

gali <- gali %>%
  filter(country  %in% countries_to_keep$`gali$country`)

#Number of ventures in main sample 
#kable(countries_to_keep)
```

#GEM NES Data

```{r}
gem_nes <- import("Data/GEM-NES.csv")

#Clean and tidy gem NES
gem_nes <- clean_names(gem_nes)
gem_nes <- gem_nes %>%
  row_to_names(row_number = 2)

gem_nes[,-2] <- gem_nes[,-2] %>% mutate_if(is.character,as.numeric)

gem_nes$inst_hc_nes <- rowMeans(gem_nes[,(5:6)])
gem_nes <- clean_names(gem_nes)
gem_nes <- gem_nes %>%
  rename(country = economy) %>%
  rename(inst_fc_gem = 	financing_for_entrepreneurs)

gem_nes <- gem_nes [, -(5:6)]

#Normalize NES data (all data is based on a scale from 1 to 5)
preproc_nes <- preProcess(gem_nes[,(4:5)], method=c("range"))

gem_nes <- predict(preproc_nes, gem_nes)

```

#GEM APS data

```{r}
#gem_aps <- import("Data/GEM-APS.csv")
#Clean and tidy gem APS
#gem_aps <- clean_names(gem_aps)
#gem_aps <- gem_aps %>%
  #row_to_names(row_number = 2)
#gem_aps[,-2] <- gem_aps[,-2] %>% mutate_if(is.character,as.numeric)
#gem_aps <- clean_names(gem_aps)
#gem_aps <- gem_aps %>%
  #rename(country = economy) %>%
  #rename(inst_signaling_aps = high_status_to_successful_entrepreneurs) #%>%
  #rename(inst_hc_aps = established_business_ownership) %>%
  #rename(inst_sc_aps = total_early_stage_entrepreneurial_activity_tea)

#All APS data is a percentage. We devide by 100 to gat [0,1] range
#gem_aps$inst_signaling_gem <- gem_aps$inst_signaling_aps /100 
#gem_aps$inst_hc_aps <- gem_aps$inst_hc_aps /100 + 1
#gem_aps$inst_sc_gem <- gem_aps$inst_sc_aps /100 + 1

```

#Final GEM data

```{r}
gem <- gem_nes
#join nes and aps on country and year 
#gem <- left_join(gem_nes, gem_aps, by=c("country", "year", "code"))
#gem$inst_hc_gem <- rowMeans(gem[,c("inst_hc_nes", "inst_hc_aps")]) + 1
#gem$inst_hc_gem <- gem$inst_hc_nes

#gem <- gem %>%
  #select(code, country, year, ends_with('_gem'))

gem$country[gem$country == "United States"] <-"USA"

gem <- gem %>%
  group_by(country) %>%
  mutate(inst_fc_gem = mean(inst_fc_gem, na.rm= TRUE))
#  inst_hc_gem = mean(inst_hc_gem, na.rm= TRUE),
  #inst_sc_gem = mean(inst_sc_gem, na.rm= TRUE),
#  inst_signaling_gem = mean(inst_signaling_gem, na.rm= TRUE))

#gem <- select(gem, -year, -inst_hc_nes)

gem <- gem %>% distinct()

```

#CGI data
```{r}
gci <- import("Data/GCI_2018.xlsx")

gci <- filter(gci, `Series Global ID` == "EOSQ089" |
                   `Series Global ID` == "EOSQ425" |
                   `Series Global ID` == "EOSQ109")
gci <- filter(gci, Edition ==2018 )

gci <- filter(gci, Attribute == "VALUE")
gci <- select(gci, -c("Index", "Edition", "Freeze date", "Series units", "Series order", "Series code (if applicable)", "Attribute", "Series name"))

gci <- gci %>%
  gather(country, score, -`Series Global ID`)

gci$country[gci$country == "United States"] <-"USA"

gci <- gci %>%
  filter(country  %in% countries_to_keep$`gali$country`)

gci$score <- round(as.numeric(gci$score),2)

gci <- clean_names(gci)

gci <- gci %>%
  pivot_wider(names_from = series_global_id, 
              values_from = score)

#preproc_gci <- preProcess(gci %>%
#                           select(starts_with("EOSQ"))
#                          , method=c("range"))
#gci <- predict(preproc_gci, gci)

#gci$inst_fc_gci <- rowMeans(gci[,c('EOSQ425','EOSQ089')], na.rm = TRUE)

#gci$inst_sc_gci <- gci$EOSQ109

#gci <- select(gci, !starts_with('EOSQ'))

```

#Attention to capital - venture data

```{r Cleaning data of acceleration benefit ranking }

#Removing ventures that ranked benefits outside of given scale (i.e. 1 to 7)

gali <- filter(gali, accel_ben_rank_network <= 7 & accel_ben_rank_business_skills
<= 7 & accel_ben_rank_mentorship <= 7 & accel_ben_rank_access_investors <= 7 & accel_ben_rank_direct_funding <= 7 & accel_ben_rank_access_entreprene <= 7 & accel_ben_rank_aware_cred <= 7 )

#Removing ventures that gave the same ranking to two or more benefits 
gali <- filter(gali, 
               accel_ben_rank_network != accel_ben_rank_business_skills &
                 accel_ben_rank_network != accel_ben_rank_mentorship & 
                 accel_ben_rank_network != accel_ben_rank_access_investors &
                 accel_ben_rank_network != accel_ben_rank_direct_funding &
                 accel_ben_rank_network != accel_ben_rank_access_entreprene &
                 accel_ben_rank_network != accel_ben_rank_aware_cred )

gali <- filter(gali, 
                 accel_ben_rank_business_skills != accel_ben_rank_mentorship & 
              accel_ben_rank_business_skills != accel_ben_rank_access_investors &
                 accel_ben_rank_business_skills != accel_ben_rank_direct_funding &
              accel_ben_rank_business_skills != accel_ben_rank_access_entreprene &
                 accel_ben_rank_business_skills != accel_ben_rank_aware_cred )

gali <- filter(gali, 
                accel_ben_rank_mentorship != accel_ben_rank_access_investors &
                 accel_ben_rank_mentorship != accel_ben_rank_direct_funding &
                 accel_ben_rank_mentorship != accel_ben_rank_access_entreprene &
                 accel_ben_rank_mentorship != accel_ben_rank_aware_cred )

gali <- filter(gali, 
              accel_ben_rank_access_investors != accel_ben_rank_direct_funding &
            accel_ben_rank_access_investors != accel_ben_rank_access_entreprene &
                 accel_ben_rank_access_investors != accel_ben_rank_aware_cred )

gali <- filter(gali, 
              accel_ben_rank_direct_funding != accel_ben_rank_access_entreprene &
                 accel_ben_rank_direct_funding != accel_ben_rank_aware_cred )

gali <- filter(gali, 
              accel_ben_rank_access_entreprene != accel_ben_rank_aware_cred )

```

```{r}
#Rank 1 shows the most important benefit (highest attention)

#Normalize Attention data (all data is based on a scale from 1 to 7)
preproc_gali <- preProcess(gali %>%
                            select(starts_with("accel_ben_rank_"))
                          , method=c("range"))
gali <- predict(preproc_gali, gali)

#Calculating the average of different benifits to group them into one of the 3 categories of capital plus signaling.

gali$attention_hc <- 1- round(gali$accel_ben_rank_business_skills,2)

gali$attention_signaling <- 1- round((gali$accel_ben_rank_aware_cred),2)

gali$attention_sc <- 1- round((rowMeans(gali[,
                                  c('accel_ben_rank_network',
                                'accel_ben_rank_access_entreprene',
                                'accel_ben_rank_access_investors')],
                              na.rm=TRUE)),2)

gali$attention_fc <- 1- round(gali$accel_ben_rank_direct_funding,2)

```

#Join all 3 datasets 
```{r}
gali_joined <- left_join(gali, gem, by = c("country", "year"))

gali_joined <- left_join(gali_joined, gci, by = "country")

```

#Fiancial Capital acquisiton total
```{r}
gali_joined  <- gali_joined %>% mutate(capital_raised0 = gali_joined %>% select(starts_with("inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised1 = gali_joined %>% select(starts_with("fu1inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised2 = gali_joined %>% select(starts_with("fu2inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised3 = gali_joined %>% select(starts_with("fu3inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised4 = gali_joined %>% select(starts_with("fu4inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised_tot = gali_joined %>% select(starts_with("capital_raised")) %>% 
                         rowSums(na.rm = TRUE))

```

#Human capital acquisition 
```{r}
gali_joined  <- gali_joined %>% mutate(hr_tot = gali_joined %>%
                         select(contains("_employees")) %>% 
                         rowSums(na.rm = TRUE))
```

#Social capital acquisition 
```{r}
gali_joined <- gali_joined %>% mutate(equity_sources_tot = gali_joined %>%
                         select(contains("equityfrom")) %>% 
                         rowSums(na.rm = TRUE))
```

#Revenues
```{r}
gali_joined <- gali_joined %>% mutate(revenue_tot = gali_joined %>%
                         select(contains("revenues")) %>% 
                         rowSums(na.rm = TRUE))
```

#Mixed model analysis 
```{r}
fit0 <- lmer(capital_raised_tot ~  attention_fc + (attention_fc|country), data = gali_joined, REML = F)

fit1 <- lmer(capital_raised_tot ~ attention_fc + attention_fc*EOSQ425 + (1 |country), data = gali_joined, REML = F)

fit2 <- lmer(capital_raised_tot ~ attention_fc + attention_fc*EOSQ089 + (1 |country), data = gali_joined, REML = F)

fit3 <- lmer(capital_raised_tot ~ attention_fc + attention_fc*inst_fc_gem + (1 |country), data = gali_joined, REML = F)

summary(fit2)

```

#Exploring with plots
```{r}
ggplot(gali_joined, aes(inst_hc_gem,inst_fc_gem)) +
  geom_point()

```

##Constrcuting human capital index (control variable)
###Graduate percentage, Prior C-level Executive Percentage, Average Team Tenure , Team Prior Founding

```{r Human capital index}
#Graduate percentage

gali_joined$founder1_grad = ifelse(gali_joined$found_name1_education == "Masters Degree" | 
                              gali_joined$found_name1_education =="PhD" |
                              gali_joined$found_name1_education == "Some graduate degree", 1, 0) 
gali_joined$founder1_grad[is.na(gali_joined$founder1_grad)] <- 0

gali_joined$founder2_grad = ifelse(gali_joined$found_name2_education == "Masters Degree" | 
                              gali_joined$found_name2_education =="PhD" |
                              gali_joined$found_name2_education == "Some graduate degree", 1, 0) 
gali_joined$founder2_grad[is.na(gali_joined$founder2_grad)] <- 0

gali_joined$founder3_grad = ifelse(gali_joined$found_name3_education == "Masters Degree" |
                              gali_joined$found_name3_education =="PhD" |
                              gali_joined$found_name3_education == "Some graduate degree", 1, 0)
gali_joined$founder3_grad[is.na(gali_joined$founder3_grad)] <- 0

gali_joined$nb_of_founders <- gali_joined$found_name1_present + gali_joined$found_name2_present + gali_joined$found_name3_present

gali_joined <- filter(gali_joined, nb_of_founders > 0)

gali_joined$grad_pct = ((gali_joined$founder1_grad + gali_joined$founder2_grad + gali_joined$founder3_grad) / gali_joined$nb_of_founders)

#Prior C-level Executive Percentage

gali_joined$founder1_job1_clevel = ifelse(is.na(gali_joined$found_name1_job1_role) | gali_joined$found_name1_job1_role !=  "CEO/Executive Director" , 0, 1)
gali_joined$founder1_job2_clevel = ifelse(is.na(gali_joined$found_name1_job2_role) | gali_joined$found_name1_job2_role !=  "CEO/Executive Director" , 0, 1)
gali_joined$founder2_job1_clevel = ifelse(is.na(gali_joined$found_name2_job1_role) | gali_joined$found_name2_job1_role !=  "CEO/Executive Director" , 0, 1)
gali_joined$founder2_job2_clevel = ifelse(is.na(gali_joined$found_name2_job2_role) | gali_joined$found_name2_job2_role !=  "CEO/Executive Director" , 0, 1)
gali_joined$founder3_job1_clevel = ifelse(is.na(gali_joined$found_name3_job1_role) | gali_joined$found_name3_job1_role !=  "CEO/Executive Director" , 0, 1)
gali_joined$founder3_job2_clevel = ifelse(is.na(gali_joined$found_name3_job2_role) | gali_joined$found_name3_job2_role !=  "CEO/Executive Director" , 0, 1)

gali_joined$prior_clevel_pct = ((gali_joined$founder1_job1_clevel + gali_joined$founder1_job2_clevel + gali_joined$founder2_job1_clevel + 
  gali_joined$founder2_job2_clevel + gali_joined$founder3_job1_clevel + gali_joined$founder3_job2_clevel) / (gali_joined$nb_of_founders*2))

#Average Team Tenure

gali_joined$avg_teamtenure <- rowMeans(subset(gali_joined, select = c('found_name1_job1_tenure', 'found_name1_job2_tenure', 'found_name2_job1_tenure', 'found_name2_job2_tenure', 'found_name3_job1_tenure', 'found_name3_job2_tenure')), na.rm = TRUE)

gali_joined$avg_teamtenure[is.nan(gali_joined$avg_teamtenure)] <- 0

#Team Prior Founding
gali_joined$team_prior_founding <- gali_joined$found_name1_prior_fp + gali_joined$found_name2_prior_fp + gali_joined$found_name3_prior_fp + gali_joined$found_name1_prior_other + gali_joined$found_name2_prior_other + gali_joined$found_name3_prior_other + gali_joined$found_name1_prior_np + gali_joined$found_name2_prior_np + gali_joined$found_name3_prior_np

#Final Human Capital Index

gali_joined$z_team_prior_founding <- scale(gali_joined$team_prior_founding, center = TRUE, scale = TRUE)
gali_joined$z_prior_clevel_pct <- scale(gali_joined$prior_clevel_pct, center = TRUE, scale = TRUE)
gali_joined$z_avg_teamtenure <- scale(gali_joined$avg_teamtenure, center = TRUE, scale = TRUE)
gali_joined$z_grad_pct <- scale(gali_joined$grad_pct, center = TRUE, scale = TRUE)

#Cite Bolstad et al, Bioinformatics (2003)

gali_joined$normq_team_prior_founding <- normalize.quantiles(gali_joined$z_team_prior_founding,copy=TRUE)
gali_joined$normq_prior_clevel_pct <- normalize.quantiles(gali_joined$z_prior_clevel_pct,copy=TRUE)
gali_joined$normq_avg_teamtenure <- normalize.quantiles(gali_joined$z_avg_teamtenure,copy=TRUE)
gali_joined$normq_grad_pct <- normalize.quantiles(gali_joined$z_grad_pct,copy=TRUE)

#group all into an index

gali_joined$HC_index <- (gali_joined$normq_team_prior_founding + gali_joined$normq_prior_clevel_pct + gali_joined$normq_avg_teamtenure + gali_joined$normq_grad_pct)

```

##Constrcuting social capital index (control variable)

```{r}

```

##Constrcuting financial capital index (control variable)

```{r}

```

##Gender decomposition variable

```{r Gender decomposition variable}
gali <- gali %>% drop_na(found_name1_gender)

gali$found_name2_gender[is.na(gali$found_name2_gender)] <- 0
gali$found_name3_gender[is.na(gali$found_name3_gender)] <- 0

gali$founder1_f <- ifelse(gali$found_name1_gender == "F", 1,0)
gali$founder2_f <- ifelse(gali$found_name2_gender == "F", 1,0)
gali$founder3_f <- ifelse(gali$found_name3_gender == "F", 1,0)

gali$female_pct <- round(((gali$founder1_f + gali$founder2_f + gali$founder3_f) / gali$nb_of_founders ), 2)

```

```{r Capital structure study variables}

##Total invesmtment
gali_0$tot_cap_m1 <- rowSums(gali_0[ , c('inv_ownmoney_m1', 'inv_outequity_m1', 'inv_totaldebt_m1', 'inv_philan_m1')],na.rm = TRUE)

gali_0$tot_cap_sincefound <- rowSums(gali_0[ , c('inv_ownmoney_sincefound', 'inv_outequity_sincefound', 'inv_totaldebt_sincefound', 'inv_philan_sincefound')],na.rm = TRUE)

gali_0$fu1_tot_cap_m1 <- rowSums(gali_0[ , c('fu1-inv_ownmoney_m1', 'fu1-inv_outequity_m1', 'fu1-inv_totaldebt_m1', 'fu1-inv_philan_m1')],na.rm = TRUE)

##Percentage of each capital type 
###Equity
gali_0$pctequity_m1 <- gali_0$inv_outequity_m1 / gali_0$tot_cap_m1
gali_0$pctequity_sincefound <- gali_0$inv_outequity_sincefound / gali_0$tot_cap_sincefound
gali_0$fu1_pctequity_m1 <- gali_0$`fu1-inv_outequity_m` / gali_0$fu1_tot_cap_m1

###Debt
gali_0$pctdebt_m1 <- gali_0$inv_totaldebt_m1 / gali_0$tot_cap_m1
gali_0$pctdebt_sincefound <- gali_0$inv_totaldebt_sincefound / gali_0$tot_cap_sincefound
gali_0$fu1_pctdebt_m1 <- gali_0$`fu1-inv_totaldebt_m` / gali_0$fu1_tot_cap_m1

###Own Money
gali_0$pctownmoney_m1 <- gali_0$inv_ownmoney_m1 / gali_0$tot_cap_m1
gali_0$pctownmoney_sincefound <- gali_0$inv_ownmoney_sincefound / gali_0$tot_cap_sincefound
gali_0$fu1_pctownmoney_m1 <- gali_0$`fu1-inv_ownmoney_m` / gali_0$fu1_tot_cap_m1

###Phillantropy
gali_0$pctphilan_m1 <- gali_0$inv_philan_m1 / gali_0$tot_cap_m1
gali_0$pctphilan_sincefound <- gali_0$inv_philan_sincefound / gali_0$tot_cap_sincefound
gali_0$fu1_pctphilan_m1 <- gali_0$`fu1-inv_philan_m` / gali_0$fu1_tot_cap_m1

##Single source vs dominant source (> 90%)
###Single source
gali_0$single_source_m1 <- ifelse(gali_0$pctequity_m1 == 1 | 
                                  gali_0$pctdebt_m1 == 1|
                                  gali_0$pctownmoney_m1 == 1 |
                                  gali_0$pctphilan_m1 ==1 , 1, 0)

gali_0$single_source_sincefound <- ifelse(gali_0$pctequity_sincefound == 1 | 
                                  gali_0$pctdebt_sincefound == 1|
                                  gali_0$pctownmoney_sincefound == 1 |
                                  gali_0$pctphilan_sincefound ==1 , 1, 0)

gali_0$single_fu1_source_m1 <- ifelse(gali_0$fu1_pctequity_m1 == 1 | 
                                  gali_0$fu1_pctdebt_m1 == 1|
                                  gali_0$fu1_pctownmoney_m1 == 1 |
                                  gali_0$fu1_pctphilan_m1 ==1 , 1, 0)

###Dominant source
gali_0$dominant_source_m1 <- ifelse((gali_0$pctequity_m1 >= .9 | 
                                  gali_0$pctdebt_m1 >= .9 |
                                  gali_0$pctownmoney_m1 >= .9 |
                                  gali_0$pctphilan_m1 >= .9) &
                                    gali_0$single_source_m1 == 0, 1, 0)

gali_0$dominant_source_sincefound <- ifelse((gali_0$pctequity_sincefound >= .9 | 
                                  gali_0$pctdebt_sincefound >= .9 |
                                  gali_0$pctownmoney_sincefound >= .9 |
                                  gali_0$pctphilan_sincefound >= .9) &
                                  gali_0$single_source_m1 == 0, 1, 0)

gali_0$dominant_fu1_source_m1 <- ifelse((gali_0$fu1_pctequity_m1 >= .9 | 
                                  gali_0$fu1_pctdebt_m1 >= .9 |
                                  gali_0$fu1_pctownmoney_m1 >= .9 |
                                  gali_0$fu1_pctphilan_m1 >= .9) &
                                  gali_0$single_source_m1 == 0, 1, 0)
```

```{r Apirations for capital structure in 12 months and 3 years }
##Total aspired  invesmtment Pre-acceleration 
gali_0$tot_plans_cap_12m <- rowSums(gali_0[ , c('inv_plans_outequity_12months', 'inv_plans_debt_12months', 'inv_plans_philan_12months')],na.rm = TRUE)

gali_0$tot_plans_cap_3y <- rowSums(gali_0[ , c('inv_plans_outequity_3years', 'inv_plans_debt_3years', 'inv_plans_philan_3years')], na.rm = TRUE)

##Total aspired  invesmtment Post-acceleration 
gali_0$fu1_tot_plans_cap_12m <- rowSums(gali_0[ , c('fu1-inv_plans_outequity_12months', 'fu1-inv_plans_debt_12months', 'fu1-inv_plans_philan_12months')], na.rm = TRUE)

gali_0$fu1_tot_plans_cap_3y <- rowSums(gali_0[ , c('fu1-inv_plans_outequity_3years', 'fu1-inv_plans_debt_3years', 'fu1-inv_plans_philan_3years')], na.rm = TRUE)

##Aspired percentage of each capital type pre-acceleration 
###Equity
gali_0$pctequity_plan12m <- gali_0$inv_plans_outequity_12months / gali_0$tot_plans_cap_12m
gali_0$pctequity_plan3y <- gali_0$inv_plans_outequity_3years / gali_0$tot_plans_cap_3y

###Debt
gali_0$pctdebt_m1 <- gali_0$inv_plans_debt_12months / gali_0$tot_plans_cap_12m
gali_0$pctdebt_sincefound <- gali_0$inv_plans_debt_3years / gali_0$tot_plans_cap_3y

###Phillantropy
gali_0$pctphilan_m1 <- gali_0$inv_plans_philan_12months / gali_0$tot_plans_cap_12m
gali_0$pctphilan_sincefound <- gali_0$inv_plans_philan_3years / gali_0$tot_plans_cap_3y

##Aspired percentage of each capital type post-acceleration 
###Equity
gali_0$fu1_pctequity_plan12m <- gali_0$`fu1-inv_plans_outequity_12months` / gali_0$fu1_tot_plans_cap_12m
gali_0$fu1_pctequity_plan3y <- gali_0$`fu1-inv_plans_outequity_3years` / gali_0$fu1_tot_plans_cap_3y

###Debt
gali_0$fu1_pctdebt_plan12m <- gali_0$`fu1-inv_plans_debt_12months` / gali_0$fu1_tot_plans_cap_12m
gali_0$fu1_pctdebt_plan3y <- gali_0$`fu1-inv_plans_debt_3years` / gali_0$fu1_tot_plans_cap_3y

###Phillantropy
gali_0$fu1_pctphilan_plan12m <- gali_0$`fu1-inv_plans_philan_12months` / gali_0$fu1_tot_plans_cap_12m
gali_0$fu1_pctphilan_plan3y <- gali_0$`fu1-inv_plans_philan_3years` / gali_0$fu1_tot_plans_cap_3y

```