### 2021-02-25 Manuscript.Rmd

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "Ouafaa Hmaddi"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "ohmaddi@uoregon.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "University of Oregon"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.
```{r message=FALSE, warning=FALSE, include=FALSE}
library(rio)
library(tidyverse)
library(janitor)
library(ggplot2)
library(psych)
library(knitr)

```

```{r Countries to analyse in main sample, message=FALSE, warning=FALSE, include=FALSE}
gali <- import("Data/GALI_External_DataRelease_2020.xlsx")

count(gali,info_venture_country_hq)

#Removing countries with less than 6 ventures.

countries_to_keep <- gali %>% 
  group_by(gali$info_venture_country_hq) %>%
  summarize(Count=n()) %>%
  mutate(Percent = round((Count/sum(Count)*100))) %>%
  arrange(desc(Count))

countries_to_keep <- filter(countries_to_keep, `gali$info_venture_country_hq` != 'NA')

countries_to_keep <- filter(countries_to_keep, Count >9)
countries_to_keep$`gali$info_venture_country_hq`

gali <- gali %>%
  filter(info_venture_country_hq  %in% countries_to_keep$`gali$info_venture_country_hq`)

```

```{r message=FALSE, warning=FALSE, include=FALSE}
gali <- filter(gali, found_name1_age > 14 | is.na(found_name1_age))
gali <- filter(gali, found_name2_age > 14 | is.na(found_name2_age))
gali <- filter(gali, found_name3_age > 14 | is.na(found_name3_age))
gali <- filter(gali, found_name1_age < 70 | is.na(found_name1_age))
gali <- filter(gali, found_name2_age < 70 | is.na(found_name2_age))
gali <- filter(gali, found_name3_age < 70 | is.na(found_name3_age))

#Keep only for porfit firms 
gali <- filter(gali, info_legal_status == "For-profit company" )

#firms that are 5 years or younger when they entered the acceleration program (when data was collected)

if (gali$application_year == 2013) { gali <- filter(gali, info_founding_year > 2007)
}  else if (gali$application_year == 2014) {gali <- filter(gali, info_founding_year > 2008)
} else if (gali$application_year == 2015) {gali <- filter(gali, info_founding_year > 2009)
}  else if (gali$application_year == 2016) {gali <- filter(gali, info_founding_year > 2010)
} else if (gali$application_year == 2017) {gali <- filter(gali, info_founding_year > 2011)
}  else gali <- filter(gali, info_founding_year > 2012)
```

```{r Countries to keep in sample with valid data, message=FALSE, warning=FALSE, include=FALSE}

gali <- gali %>%
  rename(country = info_venture_country_hq) %>%
  rename(year = program_year)

gali$country[gali$country == "United States of America"] <-"USA"
gali$country[gali$country == "Tanzania, United Republic of"] <-"Tanzania"

countries_to_keep <- gali %>% 
  group_by(gali$country) %>%
  summarize(Count=n()) %>%
  mutate(Percent = round((Count/sum(Count)*100))) %>%
  arrange(desc(Count))

countries_to_keep <- filter(countries_to_keep, `gali$country` != 'NA')

countries_to_keep <- filter(countries_to_keep, Count >9)
countries_to_keep$`gali$country`

gali <- gali %>%
  filter(country  %in% countries_to_keep$`gali$country`)

```

#CGI data
```{r message=FALSE, warning=FALSE, include=FALSE}
gci <- import("Data/GCI_2018.xlsx")

gci <- filter(gci, `Series Global ID` == "EOSQ089" |
                   `Series Global ID` == "EOSQ425" |
                   `Series Global ID` == "EOSQ109" |
                   `Series Global ID` == "DOMCREDITGDP" |
                   `Series Global ID` == "EOSQ403")
gci <- filter(gci, Edition ==2018 )

gci <- filter(gci, Attribute == "VALUE")
gci <- select(gci, -c("Index", "Edition", "Freeze date", "Series units", "Series order", "Series code (if applicable)", "Attribute", "Series name"))

gci <- gci %>%
  gather(country, score, -`Series Global ID`)

gci$country[gci$country == "United States"] <-"USA"

gci <- gci %>%
  filter(country  %in% countries_to_keep$`gali$country`)

gci$score <- round(as.numeric(gci$score),2)

gci <- clean_names(gci)

gci <- gci %>%
  pivot_wider(names_from = series_global_id, 
              values_from = score)
```

#Plotting all countries along the 3 dimensions

```{r echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}

library(plot3D)

with(gci, text3D(EOSQ403, EOSQ425, EOSQ109, 
  labels = gci$country,  
  col = gg.col(100), theta = 60, phi = 20,
  xlab = "Human capital", ylab = "Financail capital", zlab = "Social capital", 
  main = "Institutional level resources", cex = 0.6, 
  bty = "g", ticktype = "detailed", d = 2,
  clab = c("Urban","Pop"), adj = 1, font = 1.5))

#Zoom into to the middle for more clarity 
plotdev(xlim = c(3, 4), ylim = c(3, 5.5), 
         zlim = c(3, 3.8))

```

#Attention to capital - venture data

```{r Cleaning data of acceleration benefit ranking, message=FALSE, warning=FALSE, include=FALSE}

#Removing ventures that ranked benefits outside of given scale (i.e. 1 to 7)

gali <- filter(gali, accel_ben_rank_network <= 7 & accel_ben_rank_business_skills
<= 7 & accel_ben_rank_mentorship <= 7 & accel_ben_rank_access_investors <= 7 & accel_ben_rank_direct_funding <= 7 & accel_ben_rank_access_entreprene <= 7 & accel_ben_rank_aware_cred <= 7 )

#Removing ventures that gave the same ranking to two or more benefits 
gali <- filter(gali, 
               accel_ben_rank_network != accel_ben_rank_business_skills &
                 accel_ben_rank_network != accel_ben_rank_mentorship & 
                 accel_ben_rank_network != accel_ben_rank_access_investors &
                 accel_ben_rank_network != accel_ben_rank_direct_funding &
                 accel_ben_rank_network != accel_ben_rank_access_entreprene &
                 accel_ben_rank_network != accel_ben_rank_aware_cred )

gali <- filter(gali, 
                 accel_ben_rank_business_skills != accel_ben_rank_mentorship & 
              accel_ben_rank_business_skills != accel_ben_rank_access_investors &
                 accel_ben_rank_business_skills != accel_ben_rank_direct_funding &
              accel_ben_rank_business_skills != accel_ben_rank_access_entreprene &
                 accel_ben_rank_business_skills != accel_ben_rank_aware_cred )

gali <- filter(gali, 
                accel_ben_rank_mentorship != accel_ben_rank_access_investors &
                 accel_ben_rank_mentorship != accel_ben_rank_direct_funding &
                 accel_ben_rank_mentorship != accel_ben_rank_access_entreprene &
                 accel_ben_rank_mentorship != accel_ben_rank_aware_cred )

gali <- filter(gali, 
              accel_ben_rank_access_investors != accel_ben_rank_direct_funding &
            accel_ben_rank_access_investors != accel_ben_rank_access_entreprene &
                 accel_ben_rank_access_investors != accel_ben_rank_aware_cred )

gali <- filter(gali, 
              accel_ben_rank_direct_funding != accel_ben_rank_access_entreprene &
                 accel_ben_rank_direct_funding != accel_ben_rank_aware_cred )

gali <- filter(gali, 
              accel_ben_rank_access_entreprene != accel_ben_rank_aware_cred )

```

#Reverse code attention variable 
```{r message=FALSE, warning=FALSE, include=FALSE}
#Reverse code fundtion 
reverse <- function(x){x*-1- +8}

gali_att <- as.data.frame(lapply(gali %>%
                            select(starts_with("accel_ben_rank_")), reverse))

gali <- cbind(gali %>%
        select(-starts_with("accel_ben_rank_")), gali_att)

```

```{r message=FALSE, warning=FALSE, include=FALSE}

mean_norm_minmax <- function(x){
                                (x- mean(x)) /(max(x)-min(x))
                               }

mean_norm_att <- as.data.frame(lapply(gali %>%
                            select(starts_with("accel_ben_rank_")), mean_norm_minmax))

gali <- cbind(gali %>%
        select(-starts_with("accel_ben_rank_")), mean_norm_att)
```

#Join all 3 datasets 
```{r message=FALSE, warning=FALSE, include=FALSE}

gali_joined <- left_join(gali, gci, by = "country")

```

#Fiancial Capital acquisiton total
```{r message=FALSE, warning=FALSE, include=FALSE}
gali_joined  <- gali_joined %>% mutate(capital_raised0 = gali_joined %>% select(starts_with("inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised1 = gali_joined %>% select(starts_with("fu1inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised2 = gali_joined %>% select(starts_with("fu2inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised3 = gali_joined %>% select(starts_with("fu3inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised4 = gali_joined %>% select(starts_with("fu4inv_") & ends_with("_m1")) %>% 
                         rowSums(na.rm = TRUE))

gali_joined  <- gali_joined %>% mutate(capital_raised_tot = gali_joined %>% select(starts_with("capital_raised")) %>% 
                         rowSums(na.rm = TRUE))

```

#Revenues
```{r message=FALSE, warning=FALSE, include=FALSE}
gali_joined <- gali_joined %>% mutate(revenue_tot = gali_joined %>%
                         select(contains("revenues")) %>% 
                         rowSums(na.rm = TRUE))
```

#Human capital acquisition 
```{r message=FALSE, warning=FALSE, include=FALSE}
gali_joined  <- gali_joined %>% mutate(hr_tot = gali_joined %>%
                         select(contains("_employees")) %>% 
                         rowSums(na.rm = TRUE))
```

#Social capital acquisition 
```{r message=FALSE, warning=FALSE, include=FALSE}
gali_joined <- gali_joined %>% mutate(equity_sources_tot = gali_joined %>%
                         select(contains("equityfrom")) %>% 
                         rowSums(na.rm = TRUE))
```

# Results

```{r message=FALSE, warning=FALSE, include=FALSE}
#Function to extract coefficients of summary table

summary_table <- function(x){ x$coefficients %>%
  as.data.frame %>%
    rename(
      "$b$" = "Estimate"
    , "SE" = "Std. Error"
    , "$t$" = "t value"
  ) %>%
  mutate(
    Effect = papaja:::prettify_terms(rownames(.))
  ) %>%
  printnum(
    digits = c(2, 2, 2, 3)
    , gt1 = c(TRUE, TRUE, TRUE, TRUE, FALSE, FALSE)
    , zero = c(TRUE, TRUE, TRUE, TRUE, FALSE, FALSE)
  ) %>%
  select(Effect, `$b$`, `SE`, `$t$`)}

#for glmer

summary_table_z <- function(x){ x$coefficients %>%
  as.data.frame %>%
    rename(
      "$b$" = "Estimate"
    , "SE" = "Std. Error"
    , "$z$" = "z value"
  ) %>%
  mutate(
    Effect = papaja:::prettify_terms(rownames(.))
  ) %>%
  printnum(
    digits = c(2, 2, 2, 3)
    , gt1 = c(TRUE, TRUE, TRUE, TRUE, FALSE, FALSE)
    , zero = c(TRUE, TRUE, TRUE, TRUE, FALSE, FALSE)
  ) %>%
  select(Effect, `$b$`, `SE`, `$z$`)}

```

#Mixed model analysis - Human capital
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(lme4)
fit0hc <- glmer(participated ~  accel_ben_rank_business_skills + (accel_ben_rank_business_skills|country), data = gali_joined, family=binomial)

fit1hc <- glmer(participated ~  accel_ben_rank_business_skills*EOSQ403 + (accel_ben_rank_business_skills|country), data = gali_joined, family=binomial)

fit00hc <- lmer(hr_tot ~  accel_ben_rank_business_skills + (accel_ben_rank_business_skills|country), data = gali_joined, REML = F)

fit11hc <- lmer(hr_tot ~  accel_ben_rank_business_skills*EOSQ403 + (accel_ben_rank_business_skills|country), data = gali_joined, REML = F)
hc0 <- summary_table_z(summary(fit0hc))
hc1 <- summary_table_z(summary(fit1hc))
hc00 <- summary_table(summary(fit00hc))
hc11 <- summary_table(summary(fit11hc))

rownames(hc0) <- c("Intercept","Attention to Human capital(AHC)")
rownames(hc1) <- c("Intercept","AHC", "Ease of finding skilled employees", "AHCxEase of finding skilled employees")
rownames(hc00) <- c("Intercept","Attention to Human capital")
rownames(hc11) <- c("Intercept","AHC", "Ease of finding skilled employees", "AHCxEase of finding skilled employees")

kable(hc0[,-1]) 
kable(hc1[,-1])
kable(hc00[,-1])
kable(hc11[,-1])

```

#Mixed model analysis - Financial capital
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(lme4)
fit0fc <- glmer(participated ~  accel_ben_rank_direct_funding + (accel_ben_rank_direct_funding|country), data = gali_joined, family=binomial)

fit1fc <- glmer(participated ~ accel_ben_rank_direct_funding*EOSQ425 + (accel_ben_rank_direct_funding |country ), data = gali_joined, family=binomial)

fit2fc <- glmer(participated ~ accel_ben_rank_direct_funding*EOSQ089 + (accel_ben_rank_direct_funding |country ), data = gali_joined, family=binomial)

fit3fc <- glmer(participated ~ accel_ben_rank_direct_funding*DOMCREDITGDP + (accel_ben_rank_direct_funding |country), data = gali_joined, family=binomial)

summary(fit0fc)
summary(fit1fc)
summary(fit2fc)
summary(fit3fc)

fit00fc <- lmer(capital_raised_tot ~   + (accel_ben_rank_direct_funding|country), data = gali_joined, REML = F)

fit11fc <- lmer(capital_raised_tot ~ accel_ben_rank_direct_funding*EOSQ425 + (accel_ben_rank_direct_funding |country ), data = gali_joined, REML = F)

fit22fc <- lmer(capital_raised_tot ~ accel_ben_rank_direct_funding*EOSQ089 + (accel_ben_rank_direct_funding |country ), data = gali_joined, REML = F)

fit33fc <- lmer(capital_raised_tot ~ accel_ben_rank_direct_funding*DOMCREDITGDP + (accel_ben_rank_direct_funding |country), data = gali_joined, REML = F)

summary(fit00fc)
summary(fit11fc)
summary(fit22fc)
summary(fit33fc)

```

#Mixed model analysis - Social capital

```{r echo=FALSE, message=FALSE, warning=FALSE}
fit0sc <- lmer(participated ~ accel_ben_rank_network + (accel_ben_rank_network |country), data = gali_joined, REML = F)

fit1sc <- lmer(participated ~ accel_ben_rank_network*EOSQ109 + (accel_ben_rank_network | country), data = gali_joined, REML = F)

coef(summary(fit0sc))
coef(summary(fit1sc))

fit00sc <- lmer(equity_sources_tot ~ accel_ben_rank_network + (accel_ben_rank_network |country), data = gali_joined, REML = F)

fit11sc <- lmer(equity_sources_tot ~ accel_ben_rank_network*EOSQ109 + (accel_ben_rank_network | country), data = gali_joined, REML = F)

coef(summary(fit00sc))
coef(summary(fit11sc))

```

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup